---
title: "Data Challenge 1"
author: "Yifan Lu"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

## load in the required libraries 

```{r}
library(readr)
library(tidyverse)
library(here)

## read in the data using here() function
state_milk_production <- read_csv(here("data/state_milk_production.csv"))
clean_cheese <- read_csv(here("data/clean_cheese.csv"))
fluid_milk_sales <- read_csv(here("data/fluid_milk_sales.csv"))
milk_products_facts <- read_csv(here("data/milk_products_facts.csv"))
milkcow_facts <- read_csv(here("data/milkcow_facts.csv"))

## disable scientific notation
options(scipen = 999)
```
# Figure 1
Make a scatter plot showing the milk production in pounds versus the average number of milk cows using milkcow_facts. Write a caption for this plot.

The x-axis will be the number of milk cows (in millions) and the y-axis will be milk production (in millions of pounds)
Color the plot using a continuous variable of years
Label the axes (do not use the variable name)
Add a title for the plot
Label the legend (do not use the variable name)
```{r}
# Convert milk production from pounds to millions of pounds
milkcow_facts$milk_production_lbs_millions <- milkcow_facts$milk_production_lbs / 1e6
# Convert the number of milk cows to millions
milkcow_facts$avg_milk_cow_number_millions <- milkcow_facts$avg_milk_cow_number / 1e6

# Plotting the data
milkcow_facts %>%
    ggplot(aes(
        x = avg_milk_cow_number_millions, # Set x-axis to the number of milk cows (in millions)
        y = milk_production_lbs_millions, # Set y-axis to milk production (in millions of pounds)
        color = year # Color the plot using a continuous variable of years 
    )) +  
    geom_point() + # Plot the data points in the graph
    scale_color_continuous(name = "Year") + # Color the year continuously
    labs(
        title = "Milk Production vs Number of Milk Cows", # Set the title of the plot
        x = "Number of Milk Cows (in millions)", # Label for x-axis
        y = "Milk Production (in millions of pounds)", # Label for y-axis
        caption = "Figure 1: This scatter plot visualizes the relationship between the number of milk cows\n and milk production over time. Each point represents a year, with its position\n determined by the number of milk cows (x-axis) and total milk production (y-axis) \n for that year. The color of each point corresponds to the specific year." # Caption for the plot
        ) +
    theme(plot.caption = element_text(hjust = 0.5)) # Adjust the caption to middle

```

# Figure 2
Examine the milk total milk production in each region over time using state_milk_production. We will plot line plots for the milk production versus the year faceted by region. Write a caption for this plot.

First, find the total production of milk each year by region (hint: use summarise() to find this)
The x-axis will be time and the y-axis will be total milk produced (in millions of pounds) in the region that year
Make line plots and color each line a different color
Facet the plot by region
Label the axes (do not use the variable name)
Add a title for the plot
Remove the legend from the plot
```{r}
# Group the data by year and region
state_milk_production %>%
    group_by(year, region) %>%
    # Summarise the total milk production by region in millions of pounds
    summarise(total_production_by_region_millions = sum(milk_produced)/ 1e6) %>%
    # Plot the data
    ggplot(aes(
        x = year, # Set x-axis to year
        y = total_production_by_region_millions, # Set y-axis to total milk production by region in millions of pounds
        color = region # Color the plot by region
    )) +
    geom_line() + # Plot the data as a line graph
    facet_wrap(~region) + # Create a separate plot for each region
    labs(
        x = "Year", # Label for x-axis
        y = "Total Milk Production (in millions of pounds)", # Label for y-axis
        title = "Total Milk Production Over Time by Region", # Set the title of the plot
        caption = "Figure 2: This line graph visualizes the total milk production over time by region.\n Each line represents a region, with its position determined by the year (x-axis) \nand the total milk production in that region (y-axis). The color of each line corresponds \nto a specific region. The graph is faceted by region, meaning there is a separate \nplot for each region. This allows for easy comparison of milk production trends across regions." # Caption for the plot
    ) +
    theme(
        legend.position = "none", # Hide the legend
        plot.caption = element_text(hjust = 0.5) # Adjust the caption to middle
    )
```

# Figure 3
Make a line plot showing the amount (in pounds) of each milk products sold over time using fluid_milk_sales. Write a caption for this plot.

Do not include the category ‘Total Production’ in the plot
The x-axis will be time and the y-axis will be pounds of product
Each milk product will be a separate line on the plot
A different color will be assigned for each milk product
Make the plot on the log scale for better visualization
Label the axes (do not use the variable name)
Add a title for the plot
Label the legend (do not use the variable name)

```{r}
fluid_milk_sales %>% 
    filter(milk_type != "Total Production") %>% # Exclude "Total Production" category
    group_by(milk_type) %>%
    ggplot(aes(x = year, y = pounds, color = milk_type)) + # Plot year on x-axis and pounds on y-axis, color by milk type
    geom_line(lwd = 1) + # Use line plot
    coord_trans(y = "log10") + # Transform y-axis to log scale
    labs(title = "Milk Product Sales Over Time", # Title for the plot
         x = "Year", # Label for x-axis
         y = "Pounds of Product (Log Scale)", # Label for y-axis
         color = "Milk Product", # Label for the legend
         caption = "Figure 3: This plot shows the amount (in pounds) of each milk product sold over time.\n Each milk product is represented by a separate line on the plot.\n The plot is on a log scale for better visualization") +
    theme(plot.caption = element_text(hjust = 0.5)) # Adjust the caption to middle
```

# Figure 4
Make a stacked bar plot of the different type of cheeses consumed over time using a long format version of clean_cheese that you create. Write a caption for this plot.

Data wrangling (hint: create a new dataframe called clean_cheese_long):

Remove all of the total cheese variables (we do not want these in the plot)
Change the clean_cheese data to long format using pivot_longer()
Make sure all of the cheese categories are capitalized
Plotting:

The x-axis will be time and the y-axis will be pounds of cheese
Color the stacked bar plot by the different types of cheese
Label the axes (do not use the variable name)
Add a title for the plot
Change the color of the plot using a palette from RColorBrewer (note that not all palettes will have enough colors for the plot)

```{r}
clean_cheese_long <- clean_cheese %>%
    select(-matches("Total")) %>%  # Remove all of the total cheese variables
    pivot_longer(-Year, names_to = "Cheese_Categories", values_to = "Pounds") %>% # Convert the data to longer format
    mutate("Cheese_Categories" = str_to_title(.$Cheese_Categories)) # Capitalize cheese categories

clean_cheese_long %>%
    ggplot(aes(x = Year, y = Pounds, fill = Cheese_Categories)) + # Plot year on x-axis and pounds on y-axis, color by cheese categories
    geom_bar(stat = "identity") + # Setting "stat" parameter to let height of each bar represent value in the data
    scale_fill_brewer(palette = "Set3") +
    labs(title = "Cheese Consumption Over Time",
         x = "Year", # Label for x-axis
         y = "Pounds of Cheese", # Label for y-axis
         fill = "Cheese Categories",
         caption = "Figure 4: This stacked bar plot shows how the comsumption of different categories of \n cheese changes over time. Each bar represents a year. And each sub-bars \n are colored according to their cheese categories.")  +
    theme(plot.caption = element_text(hjust = 0.5)) # Adjust the caption to middle
```

# Figure 5
Time to be creative! Make an original figure using the data. Join two of the dataframes together and use variables from both sides of the join to make this figure. This figure should tell a story and be easy to understand (just as the figures above are!). Change the colors of this plot from the default ggplot2 colors. Write a caption for this plot.

```{r}
total_milk_sales <- fluid_milk_sales %>%
    filter(milk_type == "Total Production", year >= 1980 & year <= 2014) %>%
    mutate(pounds = scale(pounds)) %>%
    select(year, pounds)

avg_milk_cow_number <- milkcow_facts %>%
    mutate(avg_milk_cow_number = scale(avg_milk_cow_number)) %>%
    select(year, avg_milk_cow_number)

milk_production_lbs <- milkcow_facts %>%
    mutate(milk_production_lbs = scale(milk_production_lbs)) %>%
    select(year, milk_production_lbs)

# Combine the three datasets
combined_data <- total_milk_sales %>%
    inner_join(avg_milk_cow_number, by = "year") %>%
    inner_join(milk_production_lbs, by = "year")

# Convert to longer format
combined_data_long <- combined_data %>%
    pivot_longer(cols = -year, names_to = "variable", values_to = "value")

# Create the plot
ggplot(combined_data_long, aes(x = year, y = value, color = variable)) +
    geom_line(lwd = 2) +
    labs(x = "Year", y = "Normalized Value", color = "Variable",
         title = "Combined Plot of Normalized Total Milk Sales, \n Average Milk Cow Number and Milk Production") +
    scale_color_brewer(palette = "Dark2") +
    scale_color_discrete(name = "Variable",
                         labels = c("Average Milk Cow Number", "Milk Production", "Total Milk Sales"))
```

