---
title: "Data Challenge 1"
author: "Yifan Lu"
output:
  html_document:
    fig_height: 6
    fig_width: 10
  word_document: default
  pdf_document: default
---
## Github Repository Link:
https://github.com/cclyfblink/Data_Challenge_1_Yifan_Lu

## load in the required libraries 

```{r}
library(readr)
library(tidyverse)
library(here)

## read in the data using here() function
state_milk_production <- read_csv(here("data/state_milk_production.csv"))
clean_cheese <- read_csv(here("data/clean_cheese.csv"))
fluid_milk_sales <- read_csv(here("data/fluid_milk_sales.csv"))
milk_products_facts <- read_csv(here("data/milk_products_facts.csv"))
milkcow_facts <- read_csv(here("data/milkcow_facts.csv"))

## disable scientific notation
options(scipen = 999)
```
# Figure 1
Make a scatter plot showing the milk production in pounds versus the average number of milk cows using milkcow_facts. Write a caption for this plot.

The x-axis will be the number of milk cows (in millions) and the y-axis will be milk production (in millions of pounds)
Color the plot using a continuous variable of years
Label the axes (do not use the variable name)
Add a title for the plot
Label the legend (do not use the variable name)
```{r}
# Convert milk production from pounds to millions of pounds
milkcow_facts$milk_production_lbs_millions <- milkcow_facts$milk_production_lbs / 1e6
# Convert the number of milk cows to millions
milkcow_facts$avg_milk_cow_number_millions <- milkcow_facts$avg_milk_cow_number / 1e6

# Plotting the data
milkcow_facts %>%
    ggplot(aes(
        x = avg_milk_cow_number_millions, # Set x-axis to the number of milk cows (in millions)
        y = milk_production_lbs_millions, # Set y-axis to milk production (in millions of pounds)
        color = year # Color the plot using a continuous variable of years 
    )) +  
    geom_point() + # Plot the data points in the graph
    scale_color_continuous(name = "Year") + # Color the year continuously
    labs(
        title = "Milk Production vs Number of Milk Cows", # Set the title of the plot
        x = "Number of Milk Cows (in millions)", # Label for x-axis
        y = "Milk Production (in millions of pounds)", # Label for y-axis
        caption = "Figure 1: This scatter plot visualizes the relationship between the number of milk cows\n and milk production over time. Each point represents a year, with its position\n determined by the number of milk cows (x-axis) and total milk production (y-axis) \n for that year. The color of each point corresponds to the specific year.\n The milk production and number of milk cows seems negatively related. And milk production increased over time." # Caption for the plot
        ) +
    theme(plot.caption = element_text(hjust = 0.5)) # Adjust the caption to middle

```

# Figure 2
Examine the milk total milk production in each region over time using state_milk_production. We will plot line plots for the milk production versus the year faceted by region. Write a caption for this plot.

First, find the total production of milk each year by region (hint: use summarise() to find this)
The x-axis will be time and the y-axis will be total milk produced (in millions of pounds) in the region that year
Make line plots and color each line a different color
Facet the plot by region
Label the axes (do not use the variable name)
Add a title for the plot
Remove the legend from the plot
```{r}
# Group the data by year and region
state_milk_production %>%
    group_by(year, region) %>%
    # Summarise the total milk production by region in millions of pounds
    summarise(total_production_by_region_millions = sum(milk_produced)/ 1e6) %>%
    # Plot the data
    ggplot(aes(
        x = year, # Set x-axis to year
        y = total_production_by_region_millions, # Set y-axis to total milk production by region in millions of pounds
        color = region # Color the plot by region
    )) +
    geom_line() + # Plot the data as a line graph
    facet_wrap(~region) + # Create a separate plot for each region
    labs(
        x = "Year", # Label for x-axis
        y = "Total Milk Production (in millions of pounds)", # Label for y-axis
        title = "Total Milk Production Over Time by Region", # Set the title of the plot
        caption = "Figure 2: This line graph visualizes the total milk production over time by region.\n Each line represents a region, with its position determined by the year (x-axis) \nand the total milk production in that region (y-axis). The color of each line corresponds \nto a specific region. The graph is faceted by region, meaning there is a separate \nplot for each region. This allows for easy comparison of milk production trends across regions. \n Lake States, Mountain, Pacific and Southern Plains have increased total milk production \n over time. While other regions only have subtle changes." # Caption for the plot
    ) +
    theme(
        legend.position = "none", # Hide the legend
        plot.caption = element_text(hjust = 0.5) # Adjust the caption to middle
    )
```

# Figure 3
Make a line plot showing the amount (in pounds) of each milk products sold over time using fluid_milk_sales. Write a caption for this plot.

Do not include the category ‘Total Production’ in the plot
The x-axis will be time and the y-axis will be pounds of product
Each milk product will be a separate line on the plot
A different color will be assigned for each milk product
Make the plot on the log scale for better visualization
Label the axes (do not use the variable name)
Add a title for the plot
Label the legend (do not use the variable name)

```{r}
fluid_milk_sales %>% 
    filter(milk_type != "Total Production") %>% # Exclude "Total Production" category
    group_by(milk_type) %>%
    ggplot(aes(x = year, y = pounds, color = milk_type)) + # Plot year on x-axis and pounds on y-axis, color by milk type
    geom_line(lwd = 1) + # Use line plot
    coord_trans(y = "log10") + # Transform y-axis to log scale
    labs(title = "Milk Product Sales Over Time", # Title for the plot
         x = "Year", # Label for x-axis
         y = "Pounds of Product (Log Scale)", # Label for y-axis
         color = "Milk Product", # Label for the legend
         caption = "Figure 3: This plot shows the amount (in pounds) of each milk product sold over time.\n Each milk product is represented by a separate line on the plot.\n The plot is on a log scale for better visualization. \n Whole milk, Flavored(Whole) and Buttermilk sales decreased over time. \n Reduced Fat(2%), Flavored (Not Whole), Low Fat(1%) and Eggnog all have increased sales. \n Skim and Flavored(Whole) sales oscillated over time.") +
    theme(plot.caption = element_text(hjust = 0.5)) # Adjust the caption to middle
```

# Figure 4
Make a stacked bar plot of the different type of cheeses consumed over time using a long format version of clean_cheese that you create. Write a caption for this plot.

Data wrangling (hint: create a new dataframe called clean_cheese_long):

Remove all of the total cheese variables (we do not want these in the plot)
Change the clean_cheese data to long format using pivot_longer()
Make sure all of the cheese categories are capitalized
Plotting:

The x-axis will be time and the y-axis will be pounds of cheese
Color the stacked bar plot by the different types of cheese
Label the axes (do not use the variable name)
Add a title for the plot
Change the color of the plot using a palette from RColorBrewer (note that not all palettes will have enough colors for the plot)

```{r}
clean_cheese_long <- clean_cheese %>%
    select(-matches("Total")) %>%  # Remove all of the total cheese variables
    pivot_longer(-Year, names_to = "Cheese_Categories", values_to = "Pounds") %>% # Convert the data to longer format
    mutate("Cheese_Categories" = str_to_title(.$Cheese_Categories)) # Capitalize cheese categories

clean_cheese_long %>%
    ggplot(aes(x = Year, y = Pounds, fill = Cheese_Categories)) + # Plot year on x-axis and pounds on y-axis, color by cheese categories
    geom_bar(stat = "identity") + # Setting "stat" parameter to let height of each bar represent value in the data
    scale_fill_brewer(palette = "Set3") +
    labs(title = "Cheese Consumption Over Time",
         x = "Year", # Label for x-axis
         y = "Pounds of Cheese", # Label for y-axis
         fill = "Cheese Categories",
         caption = "Figure 4: This stacked bar plot shows how the comsumption of different categories of \n cheese changes over time. Each bar represents a year. And each sub-bars \n are colored according to their cheese categories.\n Total cheese consumption increased over time. Among those categories, \n mozzarella cheese have the most increased proportion.")  +
    theme(plot.caption = element_text(hjust = 0.5)) # Adjust the caption to middle
```

# Figure 5
Time to be creative! Make an original figure using the data. Join two of the dataframes together and use variables from both sides of the join to make this figure. This figure should tell a story and be easy to understand (just as the figures above are!). Change the colors of this plot from the default ggplot2 colors. Write a caption for this plot.

```{r}
# Filter and scale the total milk sales data
scaled_total_milk_sales <- fluid_milk_sales %>%
    filter(year >= 1980 & year <= 2014, milk_type == "Total Production") %>% # Filter data for years 1980 to 2014 and for total production
    mutate(pounds = scale(pounds)) %>% # Scale the pounds column
    select(year, pounds) # Select the year and pounds columns

# Scale the average price of milk data
scaled_avg_price_milk <- data.frame(year = 1980:2014, dairy_ration = scale(milkcow_facts$dairy_ration))

# Combine the two data frames
combined_data <- merge(scaled_total_milk_sales, scaled_avg_price_milk, by = "year")

# Convert to long format
combined_data_long <- combined_data %>%
    pivot_longer(cols = c(pounds, dairy_ration), names_to = "variable", values_to = "value") # Pivot the data to long format

# Create the plot
ggplot(combined_data_long, aes(x = year, y = value, color = variable)) +
    geom_line(lwd = 2) + # Add lines to the plot and adjust thickness of the line
    labs(x = "Year", y = "Scaled Value",
         title = "Scaled Milk Cows' Dairy Ration and Average Price of Milk Over Time",
         color = "Variable",
         caption = "Figure 5: This plot visualizes the scaled values of Milk Cows' Dairy Ration and \n Average Price of Milk from 1980 to 2014. The values are scaled for better comparison. \n The Milk Cows' Dairy Ration is colored in red and the Average Price of Milk is colored in blue.  \nThe x-axis represents the year, and the y-axis represents the scaled values.\n As the milk cows' dairy ration began to rise at the year of 2006, \n average price of milk dropped subsequently few years later.") + # Add a caption to the plot
    theme(plot.caption = element_text(hjust = 0.5)) + # Adjust the caption to middle
    scale_color_manual(values=c("red", "blue"), # Change color scale to red and blue 
                       labels = c("Milk Cow's Dairy Ration", "Average Price of Milk")) 
```

